### 2017-10-27 doc/paper/abstract.tex
[ admin ] switching to easychair style

\begin{abstract}
Parser combinator libraries represent parsers as functions and,
using higher-order functions, define a DSL of combinators allowing
users to quickly put together programs capable of handling complex
recursive grammars. When moving to total functional languages such
as Agda, these programs cannot be directly ported: there is nothing
in the original definitions guaranteeing termination.

In this paper, we will introduce a `guarded' modal operator acting
on types and show how it allows us to give more precise types to
existing combinators thus guaranteeing totality. The resulting
library is available online together with various usage examples
at \url{https://github.com/gallais/agdarsec}.
\end{abstract}